---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=TRUE}
activity.data<-read.table(file="./activity.csv",header=TRUE, sep=",")
```


## What is mean total number of steps taken per day?
```{r echo=TRUE}
hist(activity.data$steps,main='Number of Steps',xlab='Steps')
```

Remove rows with NA values in steps column and calculate the mean and meedian
```{r echo=TRUE}
activity.data.nona<-na.omit(activity.data)
```

Mean:
```{r echo=TRUE}
mean(activity.data.nona$steps)
```

Median:
```{r echo=TRUE}
median(activity.data.nona$steps)
```

## What is the average daily activity pattern?
First sort the data frame by Steps
```{r echo=TRUE}
activity.data.sorted<-activity.data[order(activity.data[,3]),]
```

```{r echo=TRUE}
avg.daily.pattern<-tapply(activity.data.nona$steps, activity.data.nona$interval, mean)
plot(names(avg.daily.pattern), avg.daily.pattern, type='l')
```

Time interval with maximum steps
```{r echo=TRUE}
df<-as.data.frame(avg.daily.pattern)
time.interval.with.maxSteps<-subset(df,df$avg.daily.pattern == apply(df, 2, max))
time.interval.with.maxSteps
```

## Imputing missing values
Number of NAs:
```{r echo=TRUE}
num.na<-length(which(is.na(activity.data$steps)))
#num.na<-dim(activity.data)
num.na
```

NA handling Strategy:
Use the average value for the various 5-second intervals over the 2 month data collection period.
The data frame of such values avg.daily.pattern has already been created previously.

using the above strategy we will create a new data frame that substitutes the NA values with the average values.  

Activity Data : Before 
```{r echo=TRUE}
head(activity.data)
```

Apply NA handling strategy
```{r echo=TRUE}

for (i in 1:dim(activity.data)[1]) {
  if(is.na(activity.data[i,1])){ 
    row<-as.character(activity.data[i,3])
    activity.data[i,1]<-data.frame(avg.daily.pattern[row])[,]
  }
}
```

Note: Average values per interval/across days is as follows:
```{r echo=TRUE}
head(avg.daily.pattern)
```

Activity Data : After
```{r echo=TRUE}
head(activity.data)
```


Histogram, Mean, Media using the imputed data
## What is mean total number of steps taken per day?
```{r echo=TRUE}
hist(activity.data$steps,main='Number of Steps',xlab='Steps')
```

Mean:
```{r echo=TRUE}
mean(activity.data$steps)
```

Median:
```{r echo=TRUE}
median(activity.data$steps)
```
Answer: Surpisingly, no changes in mean and median values after imputing the dataset.

## Are there differences in activity patterns between weekdays and weekends?

Uses library(chron)
```{r echo=TRUE}
library(chron)
activity.data$weekend<-is.weekend(as.Date(activity.data[,2]))
#factor(activity.data$weekend)
```

Plot panels for weekend/weekday

```{r echo=TRUE}

mf_labeller <- function(var, value){
    value <- as.character(value)
    value[value=="FALSE"] <- "Weekday"
    value[value=="TRUE"]   <- "Weekend"
    return(value)
}

library(ggplot2)

#png(filename = "./instructions_fig/panelplot.png", width = 480, height = 480)

g<-ggplot(activity.data, aes(x=interval, y=steps), xlab='Time Interval',ylab='Steps') 

#add layers
g + geom_point(size=1)  + facet_grid(.~weekend, labeller=mf_labeller) + geom_line(colour="red")

 
```